MSRI's \textit{Mathematics of Machine Learning} Summer School\vspace{-1ex}

### 2019-07-29 MSRI ML Notes/MSRI_ML_notes.tex
Add MSRI notes

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{Workshop Notes} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

% Document-Specific Macros
% Primes and Maximals for the lazy person.
\newcommand{\p}{\frakp}
\newcommand{\m}{\frakm}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\supp}{supp}
\newcommand{\SH}{\mathcal{SH}}
\DeclareMathOperator{\DMack}{DMack}
\DeclareMathOperator{\Sing}{Sing}
\DeclareMathOperator{\SpcBal}{Spc_{Bal}}
\newcommand{\Rep}{\mathbf{Rep}\,}

%stable module category stuff
\DeclareMathOperator{\uHom}{\underline{Hom}}
\DeclareMathOperator{\uExt}{\underline{Ext}}
\DeclareMathOperator{\uEnd}{\underline{End}}

%group schemes
\newcommand{\Ga}{\bbG_a}

\begin{document}
%make the title page
\title{Workshop Notes\vspace{-1ex}}
\author{Nico Courts}
\date{June 24-28, 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These notes were the ones I took while attending the ``Triangulated Categories in Geometry and Representation Theory'' Workshop 
	at the University of Sydney during the week beginning June 23, 2019.
\end{abstract}

\section{The spectrum of the category of derived Mackey functors}
This talk was given by Beren Sanders on a joint work alongside Irakli Patchkoria and Christian Wimmer.

First we are going to talk about tensor triangulated geometry. Essentially we are looking at 
an essentially small tensor triangulated category and then we want to talk about the ``geometry''.
Analogous to the way we define the spectrum of a ring, we can define 
\begin{defn}
	The \textbf{spectrum} of a tensor-triangulated category $\calK$ is 
	\[\operatorname{Spec}{\calK}=\{\p\subsetneq \calK|\p\text{is a prime tensor triangulated ideal}\}\]
\end{defn}

Then if $X\in\calK$ is an object, then we can define the support of $X$ to be 
\[\operatorname{supp}(X)=\{\p\in\operatorname{Spec}(\calK)|x\notin\p\}\]

We define the closure of a prime to be the set of primes that are contained in it. There are some great universal properties.

\begin{lem}
	\begin{itemize}
		\item $\Spec(D(T)^c)\cong\Spec(R)$
		\item $\Spec(\Db(kG-\mathbf{mod}))\cong \Spec(H^\ast(G,k))$.
	\end{itemize}
	The second result is was generalized by (among others) Julia to group schemes.	
\end{lem}

\subsection{The Stable Homotopy Category}
This is the context for most of the talk for today. Let $\SH$ be the stable homotopy category.
There are a couple sources for objects here. We can consider $\sigma^\infty X$, the infinite suspension of a pointed space, 
or a cohomology theory $\bbE$ (e.g. singular cohomology).

$\SH$ is a tensor-triangulated category although I missed the tensor structure.

\begin{thm}[Balmer `10]
	There is a \textbf{comparison map}, a continuous map
	\[\rho_\calK:\Spec(\calK)\to \Spec(\End_\calK(1))\]
	where $1$ is the monoidal unit (since I haven't set up the macro for the bb font.)
\end{thm}
The idea here is that the object on the right is more easy to understand in general.

So for the example of $\SH^c$, we can compute that the endomorphism ring is $\bbZ$. 
Then if we look at any prime $\p\in\Spec(\bbZ)$, the fiber contains all the extraordinary cohomology theories 
that correspond to computing cohomology over characteristic $p$ fields.

\subsection{The spectrum of \texorpdfstring{$\SH(G)$}{SH(G)}}
If $G$ is a finite group, then $\SH(G)$ is the $G$-equivariant stable homotopy group.
In this case, the endomorphism ring is known as the Burnside ring $A(G)$.

Dress in `69 computed the spectrum of $A(G)$: For each $H\le G$, we ahve a ring homomorphism 
\[f^H:A(G)\to \bbZ\]
where $[X]$ is sent to $|X^H|.$ These induce embeddings:
\[(f^H)^\ast:\Spec(\bbZ)\to\Spec(A(G))\]
where $(p)\mapsto \p(H,p)$ and $(0)\mapsto \p(H,0)$.

\begin{defn}
	Let $O^p(G)$ be the smallest normal subgroup of $G$ such that $G/O^p(G)$ is a $p$-group.
\end{defn}
\begin{thm}[Dress '69]
	$\p(H,p)=\p(K,p)$ if and only if $O^p(H)\sim_G O^p(K).$
\end{thm}
\begin{rmk}
	For example if $p\nmid |G|$ then $O^p(H)=H$ for all $H\le G$. If $G$ is a$p$ group, then $O^p(H)=1$ for all $H$.
\end{rmk}

\subsection{Categorifying Dress' Work}
FOr each $H\le G$, we have the \textbf{geometric $H$-fixed points functor}
\[\Phi^H:\SH(G)^c\to \SH^c\]
that induce embeddings 
\[(\Phi^H)^\ast:\Spec(\SH^c)\hookrightarrow \Spec(\SH(G)^c)\]
where $\calC_{p,n}\mapsto \calP(H,p,n)$. Furthermore the copies of $\Spec(\SH^c)$ cover $\Spec(\SH(G)^c)$. T
\textbf{They do not overlap.}

The speaker, together with Paul Blamer, computed the spectrum of $\SH(G)^c$. The picture is interesting, but there is a topological connecting
them: In particular when $G=C_p$, $\calP(1,p,n)\not\subseteq \calP(G,p,n)$ but $\calP(1,p,n+1)$ is.

\begin{defn}
	For a finite $p$-group $G$, define the ``blue shift'' to be 
	\[\beta(G,n)\]
	which is the smallest $i$ such that $\calP(1,p,n+i)\subseteq\calP(G,p,n)$
\end{defn}
It was shown that for $p$ groups, $\beta(G,n)$ is the free rank of $G$.
\begin{rmk}
	For instance, $\beta((C_p)^r,n)=r$ and $\beta(C_{p^r},n)=1$
\end{rmk}

The bottom line here is that the spectrum of the $G$-equivariant can be understood ``locally''
as $\Spec(A(G))$.

\subsection{``Linearization'' of stable homotopy theories}
Some examples of what we're talking about: We can in some way think of the linearization of $\SH$ as being $\D(\bbZ)$.
Furthermore, the motivic derived category of $k$ could be the linearization of $\SH^{\bbA^1}(k)$.

In our case, we may want to consider the derived category of Mackey functors on $G$ as the linearization of $\SH(G)$. Mackey functors 
are functors that behave well with respect to the group/representation theory (e.g. induction and restriction).

One may think this category is the derived category of the abelian category of $G$-Mackey functors, but 
Kaledin (2008) argues that it should be this other category $\DMack(G)$. The speaker computed the spectrum of this 
category as well as developing a different construction.

One interpretation of the linearization in the earlier cases is that $\SH$ is the derived category of a circle 
and $D(\bbZ)$ is the derived category of the Eilenberg-Maclane spectrum $H\bbZ$. For our example, we can define 
$H\bbZ_G$ to be $H\bbZ$ considered as a trivial $G$-module in $\SH(G)$. Then we say 
\[\D(H\bbZ_G)=\DMack_{Kal}(G)\]
and compute the spectrum.

\begin{thm}
	The extension of scalars functor 
	\[\SH(G)\to D(H\bbZ_G)\]
	induces an embedding 
	\[\Spec(D(H\bbZ_G))\hookrightarrow \Spec(\SH(G))\]
	into the top and bottom layers. 
\end{thm}

The main takeaway her is that for $C_p$ the spectrum of the derived category of Mackey functors lies between that 
of $A(C_p)$ and of $\Spec(\SH(C_p))$ where the first step splits points into two and the second introduces this chromatic splitting.

\section{Supports and Cosupports in the Stable Module Category}
This series of talks was given by Julia Pevtsova.

\subsection{Tensor triangulated geometry for finite group schemes}
Let $A$ be a finite dimensional Hopf algebra and associate to $A$ the category $\Sing(A)$, the singularity category of $A$ which measures how non-semisimple
$A$ is. One of the reasons we ways to approach tensor-triangulated geometry is to develop what points are. As a disclaimer, Julia 
says that there might be some misunderstandings she has here, but that it has been developed by Balmer and Krause.

For some motivation, let $R$ be a commutative ring, $\p\in\Spec R$ and let $M$ be an $R$-module. We can gain local information 
by looking at $M_\p$ or $M\otimes_R k_p$.

Now ler $\calT=\D(R)$ where $\calT^c=\Dbperf(R)$. Take $Y\in\calT$ and $\p\in\Spec(R)$ and we can define $Y_\p$ or 
$Y\otimes_R^Lk_p$. Then we define 
\[\supp Y=\{\p\in\Spec R|Y\otimes_R^Lk_\p\ne 0\}.\]

This has some nice properties:
\begin{itemize}
	\item (Defection Property) $Y\simeq 0\Leftrightarrow \supp Y=\varnothing$.
	\item Invariant under $\otimes, \oplus,$ cones, and syzygies.
\end{itemize}
\begin{thm}[Neeman '92]
	The class of localizing subcategories in $\D(R)$ (here $R$ is Noetherian) are in one to one correspondence via $\supp$ to the subsets in $\Spec R$.

	Furthermore restricting to $\calT^c$, the thick subcategories in $\Dbperf(R)$ correspond to specialization closed subsets in $\Spec R$.
\end{thm}

\subsection{General Definitions and \texorpdfstring{$\SpcBal(\calT^c)$}{SpcBal(Tc)}}
\begin{defn}
	Let $T$ be a tensor triangulated category. We say that $\calC\subseteq\calT$ is \textbf{localizing} subcategory if $\calC$ is a full triangulated 
	subcategory that is closed under $\oplus$.

	$\calC\subseteq \calT^c$ is \textbf{thick} if it satisfies the same conditions.

	$\calC\subseteq\calT$ is a tensor ideal if it absorbs tensor products.
\end{defn}

\begin{rmk}
	In this talk $\calT$ is symmetric but this works in braided categories.
\end{rmk}
\begin{defn}
	The \textbf{Balmer Spectrum of $\calT^c$} is defined via a map $T^c\mapsto \SpcBal(T^c)$ where we send 
	$M\mapsto\supp_{Bal}M$.
\end{defn}

Now knowing $\SpcBal(T^c)$ is equivalent ot classifying thick tensor ideals in $T^c$. When $T^c=\Dbperf(R)$, 
and $\p\in\Spec R$, the map $R\to k_\p$ induces maps 
\[\Dbperf(R)\to \Dbperf(k_\p)=\D\mathbf{Vect}_{k_\p}\]
and thus 
\[\SpcBal \D\mathbf{Vect}_{k_\p}\to \SpcBal\Dbperf(R).\]

\subsection{Doing this in modular representation theory}
Classically we are doing modular representation theory when we have a finite group $G$ and a field $k$ where $\ch k| |G|$.
In essence this is just saying that our representation theory is with non-semisimple rings.

We can do this theory on a series of objects:
\begin{itemize}
	\item Finite groups (as above)
	\item Finite group schemes
	\item Finite supergroup schemes
	\item Small quantum groups
	\item Algebraic group $G$ and algebra $A$ over $G$. Then considering $\operatorname{Bimon}(G,A)$ is one possibility.
	\item Finite dimension pointed Hopf algebras
	\item Lie superalgebras (characteristic zero).
	\item One can continue the list. The idea here is that you have a big enough endomorphism ring.
\end{itemize}

Today we are focusing on finite group schemes (FGS) although we might get to say something about supergroup schemes and quantum groups.

\subsection{Affine Group Schemes}
\begin{defn}
	An \textbf{affine groups scheme $G$ over $k$} is a representable functor $G$ from commutative $k$-algebras to groups.
\end{defn}

\begin{rmk}
	$G(R)\cong \Hom_{\Algk}(k[G],R)$ is how the representation works. $k[G]$ is a commutative Hopf algebra. $G$ is finite if $\dim_kk[G]<\infty$.
\end{rmk}

\begin{defn}
	\[kG:=k[G]^\vee=\Hom_k(k[G],k)\]
	is the \textbf{group algebra} corresponding to $G$. This is a finite dimension cocommutative Hopf algebra.
\end{defn}

\begin{rmk}
	There is an equivalence of categories between finite group schemes and finite dimensional cocommutative Hopf algebras using $G\mapsto kG$. 
	Furthermore, the category $\mathbf{Rep}_kG$ is equivalent to $kG$-$\mathbf{Mod}$ (here uppercase will denote non-finite-dimensional), which we will call $G$-$\mathbf{Mod}$.
\end{rmk}

Then we can define the cohomology for $G$ via the Hochschild cohomology
\[\Ext_G^\ast(k,k)=H^\ast(G,k)=H^\ast(kG,k)\]

Some examples of these are finite groups, restricted Lie algebras (Lie algebra over fields of positive characteristic that also have a $p^{th}$ power operation). 
To construct a restricted Lie algebra, One cal let $\calG$ be some algebraic group. Then $\operatorname{Lie}\calG=\frakg$ is a restricted Lie algebra.
The representation theory of such a $\frakg$ is precisely the representation theory of the (finite-dimensional, cocommutative) Hopf algebra $u(\frakg)$,
the quotient of the universal enveloping algebra that identifies $p^{th}$ powers (in the PBW basis) with the internal $p^{th}$ powers in $\frakg$.

More generally one can consider \textbf{Frobenius kernels}. Let $F:\calG\to\calG$ be the Frobenius map. Then 
$\calG_{(1)}=\ker F$ and more generally $\calG_{(r)}=\ker F^(r)$, the $r^{th}$ iteration of $F$. These are connected
finite group schemes. These things have only a single point! However the group algebra and coordinate algebra really still capture 
all the representation theory of the original group.

If $\calG$ is an algebraic group, then we can compute $\calG_{(1)}$ and $\operatorname{Lie}\calG$ and 
\[k\calG_{(1)}\cong u(\operatorname{Lie}(\calG)).\]

When $\calG=\Ga$, then $k[\Ga]=k[T]$ where $\Delta T)=T\otimes 1+1\otimes T$ and 
\[{\Ga}_{(r)}(R)=\{a\in R|a^{p^r}=0\}.\]
Now $k[{\Ga}_{(r)}]=k[T]/(T^{p^r})$ and $k{\Ga}_{(r)}=k[u_0\cdots,u_{r-1}]/(u_0^p,\dots,u_{r-1}^p)$

Recall that $\StMod G$ is the category of $G$ representations where homs are considered module factoring through projective objects. This becomes a tensor triangulated category 
where the triangles comes from short exact sequences in $\Rep G$ and tensor is the usual one. The shift functor is the syzygy functor $\Omega^{-1}$ and the monoidal unit is $k$.

\begin{rmk}
	We have an exact sequence 
	\[\D(G)\to \K(\mathbf{Inj}(G))\to\StMod G\]
	and 
	\[\Dbperf(G)\to \Db(G)\to\mathbf{stmod} G\cong\Dsing(G).\]
	so in a way this measures the lack of regularity in the sense of commutative algebra.
\end{rmk}

Now here we can talk about $\uEnd^\ast(k)$ as well as 
\[\uHom^\ast(M,N)=\oplus_n\uHom(M,\Omega^{-n}N)\]
and 
\[\uEnd^\ast(k)\cong\widehat\Ext_G^\ast(k,k)=\hat H^\ast(G,k)\]
where the hat denotes Tate cohomology.

Now $\uEnd^\ast(k)$ acts on $\calT$ (on $\calT^c$) by acting on $\uHom^\ast(M,N)$.

THis leads to to Benson-Iyengar-Krause support theory where 
\[R=H^\ast(G,k)\to\hat H^\ast(G,k)\]
acts on $\calT$.

We can try setting $X=\Spec R=\Spec H^\ast(G,k)$ for the space where the support lives.

\subsection{\texorpdfstring{$\pi$}{pi}-points and local cohomological functors}
Recall last time we lookedd at $R=H^\ast(G,k)=\Ext_G^\ast(k,k)$ acting (via $\uEnd^\ast(k)$ on $\StMod G$, where $G$ was a finite group scheme.

For examples we considered $\frakg=\operatorname{Lie}\frakG$ where $\frakG$ is a reducitive algebraic group. Here 
$H^{odd}(\frakg,k)=0$ and $H^{ev}(\frakg,k)\simeq k[\calN]$ where $\calN$ are the set of nilpotents in $\frakg.$

\begin{thm}
	When $\frakg$ is a restricted Lie algebra,
	\[\Spec H^\ast(\frakg,k)=\calN^{[p]}=\{x\in\frakg|x^{[p]}=0\}.\]
\end{thm}
\begin{thm}[Suslin-Friedlander-Bendel '97]
	\[\Spec H^\ast(GL_{n(r)},k)\]
	is the variety of $r$ duples of $p$-nilpotent commuting matrices.
\end{thm}

Some facts about cohomology:
\begin{itemize}
	\item $H^\ast(G,k)$ is a graded commutative algebra.
	\item Friedlander and Suslin in '95 proved $H^\ast(G,k)$ is finite generated (has FG).
\end{itemize}

The idea to develop the support theory is to look in $X=\operatorname{Proj}(R)$ and then proceed via either $\pi$-points or via local cohomological functors (developed by Bens0n, Iyengar, and Krause).

\begin{ex}
	Another motivational example: let $\frakg$ be a restricted Lie algebra. Define $X=\operatorname{Proj}H^\ast(\frakg,k)\simeq\calN^{[p]}$
\end{ex}

\begin{rmk}
	$\frakg_a=\operatorname{Lie}\Ga$ is an algebra that is very small since $u(\frakg_a)=k[x]/x^p$. Now $\StMod k[x]/x^p$ is called a ``tt-field'', and for any $x\in\calN^{[p]}\subseteq\frakg$,
	setting $\frakg_a=\langle x\rangle$ gives us a functor 
	\[\calF_x:\StMod\frakg\to\StMod\frakg_a\]
	induced by inclusion.

	Now $\calF_x$ is a tensor triangulated functor and this functor exists for all $x$ in the nilcone (or its image in the projective closure $X$).
\end{rmk}
\begin{defn}
	Let $M\in\StMod G$. Then the support is 
	\[\supp M=\{x\in\calN^{[p]}|M|_{\langle x\rangle}=\calF_x(M)\neq 0\}\subseteq X=\operatorname{Proj}\calN^{[p]}\]
\end{defn}

The support satisfies all the good properties. (?)

if $G$ is a finite group scheme then as we have seen $kG$ is a finite dimensional cocommutative Hopf algebra.
\begin{defn}
	If $K/k$ is an extension, then 
	\[\alpha:K[t]/t^p\to KG=kG\otimes_k K\]
	is a $\pi$-point if it's flat and there is $\calU\subseteq G_K$ such that $\alpha$ factors through $K\calU$ where $\calU$ is 
	a unipotent abelian subgroup scheme of $G_K$.
\end{defn}

Given such a map $\alpha:K[t]/t^p\to KG$, we get a functor 
\[\calF_\alpha:\StMod G\to\StMod K[t]/t^p\]
where the codomain is a tt-field.
\begin{rmk}
	$\calF_\alpha$ is not a tensor functor, but 
	\[\calF_\alpha(M\otimes N)\cong 0\Leftrightarrow \calF_\alpha(M)\otimes\calF_\alpha(N)\cong 0\]
	and either of these happen if and only if $\calF_\alpha$ is zero on $M$ or $N$.
\end{rmk}
\begin{rmk}
	Note that this is a place where things break down for quantum groups. We no longer have factoring through an abelian subgroup.
\end{rmk}

Now return again to $X=\operatorname{Proj}H^\ast(G,k)$. Then given a $\pi$-point $\alpha$, we get maps 
\[H^\ast(G,k)\xrightarrow{-\otimes_R K}H^\ast(G,K)\xrightarrow{\alpha^\ast}H^\ast(K[t]/t^p,K)\cong K[x]\otimes\bigwedge(\lambda)\]
and this map gives us $P_\alpha$, which is the radical of the kernel of this map. This gives us a map $\alpha\mapsto P_\alpha\in X$.

The nice property here is that $\pi$-points realize all points on $X$:
\begin{thm}
	For all $\p\in X$, there exists an extension $K/k$ and $\pi$-point $\alpha$ such that $P_\alpha=\p$.
\end{thm}

\begin{defn}
	Let $M\in\StMod G$. Then the $\pi$ support is 
	\[\pi-\supp M=\{\p\in X|\calF_{\alpha_\p}(M)=\alpha^\ast_\p(M\otimes_k K)\ne 0\}\]
	notice that the construction here doesn't depent on choice of $\pi$ point, but that is something to show.
\end{defn}
\begin{defn}
	\[\pi-\operatorname{cosupp}(M)=\{\p\in X|\calF^{\alpha_\p}:=\alpha_\p^\ast(\Hom_k(K,M))\ne 0\}\]
	where this is computed ve the corestriction functor (this needs clarification).
\end{defn}
\begin{thm}
	$\pi$-$\supp(M\otimes_k N)=\pi$-$\supp(M)\cap\pi$-$\supp(N)$.
\end{thm}
\begin{thm}[BKIP `18]
	$\pi$-$\supp(M)=\varnothing$ if and only if $M=0$.
\end{thm}

\subsection{Local Cohomology Functors}
Let's take a moment to talk about the classical definition of support. When $R=H^\ast(G,k)=\Ext^\ast_G(k,k)$, we can act on 
$\uHom^\ast(M,M)$. When $M\in\stmod G$, we let $I_M=\operatorname{Ann}_R\uHom^\ast(M,M)$. Then $V(I_M)\subseteq X$ is the \textbf{support of 
$M$,} denoted $\supp M$.

To generalize this, we are going to let $\calT=\StMod G$ and associate to each $\p\in X$ a functor 
\[\Gamma_\p:\calT\to\calT\]
which selects the $\p$-local, $\p$-torsion objects.
\begin{defn}
	If $M\in\StMod G$, we can define the \textbf{localization map} $M\to M_\p$ where 
	\[\uHom^\ast(C,M_\p)=\uHom^\ast(C,M)_\p\]
	for each $C\in\calT^c$. (recall $\calT$ is $R$ linear so hom can be localized regularly.)
\end{defn}
\begin{defn}
	There exists a functorial map 
	\[\Delta\Gamma_{V(\p)}M\to M\to L_{V(\p)}M\]
	where the left part is universal $\p$-torsion and the right (I believe) is $\p$-torsion in the usual way (check the paper).
\end{defn}

Then $\Gamma_\p$ is found by taking the $\p$-local and $\p$-torsion operations in either order.

Now 
\[\Gamma_\p M:=\Gamma_p k\otimes_k M\]
and so
\begin{defn}
	$\supp_{BIK}M=\{p\in X:\Gamma_\p k\otimes M\ne 0\}$ and $\operatorname{cosupp}_{BIK}=\{\p\in X|\Hom_k(\Gamma_\p k,M)\ne 0\}$
\end{defn}

And then zooming out, we define 
\[\Gamma_\p(\calT)=\Gamma_\p(\StMod G)=\{M\in\calT|\supp_{BIK}(M)\subseteq\p\}\]

\begin{prop}
	\begin{itemize}
		\item $\pi$-$\supp(\Gamma_\p k)=\p$
		\item $\pi$-$\supp=\supp_{BIK}$ (formal from detection, tensor product)
	\end{itemize}
\end{prop}
\begin{cor}[Balmer, Friedlander-P]
	$\SpcBal(\stmod G)\cong\operatorname{Proj}H^\ast(G,k)$ (even on the level of schemes).
\end{cor}

\subsection{Localizing ideals in \texorpdfstring{$\StMod G$}{StMod G}}
Here we use the BIK local-global principle: to classify localizing $\otimes$ ideals, we need to know that $\Gamma_\p\calT$ are minimal. The ``wishful thinking''
here (which we'll flush out next time) is as follows: if we know $\pi$-$\operatorname{cosupp}=\operatorname{cosupp}_{BIK}$,
then we can use ``Neeman's minimality lemma'' which says that $\calC\subseteq\calT$ is minimal if for all $M$ and $N$ in $\calC$,
$\Hom^\ast_k(M,N)\ne 0$.

Next time we will see some more and a proof!

\subsection{Day 3}
Recall the context: we were letting $X=\operatorname{Proj}H^\ast(G,k)$ and $\calT=\StMod G$ and then looking at
\[\Gamma_p=\{M\in\calT|\supp M\subseteq \p\}\]

\begin{lem}
	$\m\in X$ is a closed point. Then $\Gamma_\m \calT$ is minimal.
\end{lem}
\begin{prf}
	We begin with a fact: $\m\in\pi$-$\supp M$ if and only if $\m\in \pi$-$\operatorname{cosupp} M$.

	To show that $\Gamma_\m T$ is minimal, it suffices to show that $\Hom_k(M.N)\ne 0$ for all $M$ and $N$. But 
	\[\pi\text{-}\operatorname{cosupp}\Hom_k(M,N)=\pi\text{-}\supp M\cap \pi\text{-}\operatorname{cosupp}N\]
	But since the $\pi$ support of $M$ is $\m$ and since $\m$ lies in the $\pi$-cosupport of $N$, then the intersection is nonempty.
	Thus since the cosupport is nontrivial, the module itself is nontrivial.
\end{prf}

\subsection{Reduction to closed points}
We want a Koszul object (in rep theory a Carlsson object): for every $b\in H^d(G,k)\cong\uHom^d(k,\Omega^{-d}k)$, we get an object
$k//b$ formed by the cone of $b:k\to\Omega^{-d}k$. Then for any sequence $\underline{b}=(b_1,\dots,b_n)$, we set 
\[k//\underline{b}=k//b_1\otimes\cdots\otimes k//b_n.\]

Now if we have any finite field extension $K/k$ (the idea here is we would like to take $k$ itself, but ), we can choose a $\m\in X_K=\operatorname{Proj}H^\ast(G,K)$ lying over $\p\in H^\ast(G,k)$.
Then by taking the additional generators for $\m$, we can pick a $\underline b$ such that $\m=\sqrt{\tilde\p+\underline b}$ where $\tilde\p$ is the lifting of $\p$ in the bigger ring.

Then we can consider the (Sylow?) functor for restriction of scalars $\operatorname{Res}^{G_K}_G:\StMod_K G\to \StMod G$
and look at the image of $\Gamma_\m(K//\underline b)$.
\begin{lem}
	$\Gamma_\m(K//\underline b)\downarrow_G\simeq \Gamma_\p k$
\end{lem}
\begin{cor}
	the image $\operatorname{Res}_G^{G_K}$ of $\Gamma_\m\StMod_K G$ in $\Gamma_\p\StMod_k G$ is dense.
\end{cor}
\begin{cor}
	$\Gamma_\p$ is minimal for all $\p$.
\end{cor}

\subsection{Support theory}
The following is tied closely to the following results:
\begin{prop}
	$\pi$-$\supp M=\varnothing$ if and only if $M\cong 0$.

	That is, the $\pi$ support has sufficient strength to detect things (``detection'').
\end{prop}
\begin{prop}
	For all $\p\in X$ there is an $\alpha$ with $\calF_\alpha$ realizing $\p$.
\end{prop}

I am unsure about how this fits in, but she also mentioned the detection of nilpotents in $H^\ast(G,k)$.
\begin{thm}[Quillen '71]
	If $G$ is a finite group, $g\in H^\ast(G,k)$ is nilpotent if and only if for all elements in an abelian $p$-group $E\subseteq G$,$g\downarrow_E$ is nilpotent.
\end{thm}

This means $H^\ast(E,k)\cong k[x_1,\dots,x_n]\otimes \bigwedge^\ast(\tau_1,\dots,\tau_n)$
\begin{thm}[Chouinard '76]
	if $G$ is a finite group, then $M$ is projective if any only if for all elementary abelian $E\le G$.
\end{thm}

Though these results we get that Jarrod's work on the lattice of elementary abelian subgroups of specific groups correspond directly to the lattice (modulo conjugacy and some finite 
group actions) of affine subspaces in spec of the ring.

Now if $\frakg$ is a Lie algebra (say $\frakg=\operatorname{Lie}\calG$ where $\calG$ is a reductive group), 
$H^\ast(\frakg,k)=k[\calN]$. Then fix some Borel subgroup $\calB\subseteq \calG$. Take any $\lambda\in \calG/\calB$ (which is apparently a flag variety?)
and extract from this a subalgebra $b_\lambda\subseteq\frakg$. Then extract a functor 
\[\calF_\lambda:\StMod\frakg\to\StMod b_\lambda\]
\begin{thm}
	Then if $M\in\StMod\frakg$, $M\cong 0$ if and only if $\calF_\lambda(M)=0$ for all $\lambda.$
\end{thm}
Let $G=\calG_{(r)}.$
\begin{defn}
	A \textbf{one-parameter subroup} is
	\[\bbG_{a_(r)}\to\calG_{(r)}\]
	where $k\bbG_{a_{(r)}}\cong k[u_0,\dots,u_{r-1}]/(u_0^p,\dots, u_{r-1}^p)$.
\end{defn}
\begin{thm}
	Nilpotence and projectivity are detected on one-parameter subgroups (up to scalar extension).
\end{thm}
Here we define the $\pi$-point:
\[K[t]/t^p\xrightarrow{t\mapsto u_{r-1}} k\bbG_{a_{(r)}}\to K\calG\]

\brk

But then we get an isogeny 
\[\operatorname{Mor}(\bbG_{a_{(r)}},G)\simeq\operatorname{Spec}H^\ast(G,k)\]

\begin{defn}
	A finite group scheme is elementary if 
	\[\calE=\bbG_{a_{(r)}}\times(\bbZ/p)^s\]
\end{defn}
\begin{thm}[Suslin '76]
	$\xi\in H^\ast(G,k)$ is nilpotent if any only if for all $K/k$ and for any elementary abelian subgroup $\calE\subseteq G_K$,
	the restriction $\xi_K\downarrow_\calE$ is nilpotent.
\end{thm}

\section{Smoothness and Properness}
Greg Stevenson gave this series of lectures.

\subsection{Beginning definitions}

Today we will be working up to dg categories. In what follows let $k$ be a fixed base field (although it works without being a field, this 
is a nice assumption). Recall the definition of a chain complex $\cdots\to X^i\xrightarrow{d^i} X^{i+1}\to\cdots$.

Recall that a chain morphism is a sequence of maps indexed by $\bbZ$ such that the obvious diagrams commute. Denote by $\Ch(k)$ the category of chain complexes of $k$
vector spaces. We have the \textbf{suspension} or \textbf{shift functor} in this category, denoted $\Sigma$. This is an autoequivalence of $\Ch(k)$.

Given a chain map $f:X\to Y$, we define $H^i(X)=\ker d^i/\Im d^{i-1}$. This gives us a functor $H^i:\Ch(k)\to \Vectk$. We say that $f$ is a quasi-isomorphisms if it induces
isomorphisms on cohomology.

We can form the mapping cone of $f$, $\cone f$ (think this is kind of a kernel and kind of a cokernel)
\[\cone(f)=(\Sigma X\oplus U,(\begin{smallmatrix}d_{\Sigma X} & 0\\ f & d_Y\end{smallmatrix}))\]

Recall the definition of acyclic objects. Then 
\begin{lem}
	$f:X\to Y$ is a quasi-isomorphism if and only if $\cone(f)$ is acyclic.
\end{lem}
\begin{rmk}
	This supports the idea that this is in fact a kernel and cokernel in a way.
\end{rmk}

Begin $\Ch(k)$ admits a symmetric monoidal structure from the total complex of the associated double 
complex (with some signs flipped). This defines a functor $-\otimes -$ into $\Ch(k)$.

We also have internal hom (aka the $\hom$ complex). Here 
\[\hom(X,Y)^n=\prod_{i\in\bbZ}\Hom_k(X^i,Y^{i+n})\]
with differential given by the graded commutator:
\[d(f)=d_Y^{i+n} f^i-(-1)^nf^{i+1}d_X^i\]

\begin{rmk}
	This is measuring, in a graded way, 
\end{rmk}
\begin{ex}
	$Z^0\hom(X,Y)=\ker d^0_{\hom(X,Y)}=\{\text{degree zero maps} f| fd=df\}$. Then we can compute that 
	\[H^0\hom(X,Y)\]
	which is precisely the homotopy classes of maps from $X$ to $Y$.
\end{ex}

\brk

We have a tensor hom adjunction (from here let $\Ch=\Ch(k)$)
\[\Ch(X\otimes Y,Z)\cong \Ch(X,\hom(Y,Z))\]
for all $X,Y,Z$ so we have the counit in the form of coevaluation
\[\varepsilon:X\otimes\hom(X,Y)\to Y\]
and given a $Z\in\Ch$, we can consider 
\[X\otimes\hom(X,Y)\otimes\hom(Y,Z)\xrightarrow{\varepsilon_Y\otimes 1}Y\otimes \hom(Y,Z)\xrightarrow{\varepsilon_Z} Z\]
and so via the adjunction we get a composition map 
\[\hom(X,Y)\otimes\hom(Y,Z)\xrightarrow{\circ}\hom(X,Z).\]

This composition map is associative. We also get unit maps $1_X:k\to\hom(X,X)$ determined by the image of $\id_X\in\Ch(X,X)$ in $\hom(X,X)$.
These maps act as units for $\circ$, using the fact that $k$ is the unit of $\otimes_k$.

So using $1_X$ and $\circ$, we get something that looks like a category! This is the primordial example of a dg-category.

\subsection{Examples and extensions}
Give $X\in\Ch$, we get a monoid $\hom(X,X)\in\Ch$. What does that mean? Well, we have a map $\hom(X,X)\otimes\hom(X,X)\to\hom(X,X)$
which is just composition, and a unit $1_X:k\to\hom(X,X)$. This is an example of a differential graded algebra. I.e. we have an algebra $A$ equipped with 
a degree 1 differential $d$ which is a graded derivation.
This is also an example of a dg-category.

\subsection{dg-categories}
\begin{defn}
	$\calA$ is a dg-category if it is given by a collection of objects along with a complex 
	$\calA(a,b)\in\Ch$ for each $a,b\in\calA$ as well as unit maps 
	\[1_a:k\to\calA(a,a)\]
	and composition maps 
	\[\circ:A(b,c)\otimes A(a,b)\to A(a,c)\]
	such that composition is associative and unital with respect to the unit maps.
\end{defn}
\begin{rmk}
	The slick way of saying this is that a dg-category is a category enriched over chain complexes. :)
\end{rmk}

The $\hom$ complexes make $\Ch$ into a dg-category. I will call this $\scrC$. If $A$ is a dga, consider the dg-cat $BA$, 
which is the category of a single object with hom set $A$.

If $B$ is any $k$-linear category we can view it as a dg category $\tilde B$ with $\tilde B(b,c)=B(b,c)$ in degree zero.

Given a dg category $\calA$, we can do a couple things to get a boring category: first, you can take the underlying category $Z^0\calA$ that has the same objects 
and the maps are the zeroth cocycles of $A(a,b)$ for each $a$ and $b$. For example $Z^0\scrC=\Ch$.

>>>>

%header stuff
\setlength{\headsep}{24pt}  % space between header and text
\pagestyle{fancy}     % set pagestyle for document
\lhead{``Mathematics of Machine Learning'' Notes} % put text in header (left side)
\rhead{Nico Courts} % put text in header (right side)
\cfoot{\itshape p. \thepage}
\setlength{\headheight}{15pt}
\allowdisplaybreaks

% Document-Specific Macros

\begin{document}
%make the title page
\title{MSRI's \textit{Mathematics of Machine Learning} Summer School\vspace{-1ex}}
\author{Nico Courts}
\date{July 29-Aug 9, 2019}
\maketitle

\renewcommand{\abstractname}{Introduction}
\begin{abstract}
	These notes were the ones I took while attending MSRI's ``Mathematics of Machine Learning'' summer school
	at the University of Washington during the two weeks above.
\end{abstract}

\section{Statistical Learning}
This series of lectures was given by Robert Shapire from Microsoft Research.
